\documentclass[12pt,oneside]{book}

\usepackage{fullpage,fancyhdr}
\usepackage{setspace}
\usepackage[colorlinks]{hyperref}
\onehalfspace

\setlength{\parindent}{0pt}

\begin{document}

\frontmatter

The Project Gutenberg EBook of Beethoven's Letters 1790-1826 Vol.~2, by Lady Wallace \par 
\vspace{12pt}
This eBook is for the use of anyone anywhere at no cost and with almost no restrictions whatsoever.
You may copy it, give it away or re-use it under the terms of the Project Gutenberg License included with
this eBook or online at www.gutenberg.net \par 
\vspace{12pt}

Produced by Juliet Sutherland, John Williams and the Online Distributed Proofreading Team.
\pagebreak

\title{Beethoven's Letters\\Volume II}
\author{Lady Wallace}
\date{1790-1826}

\maketitle

From the collection of Dr.~Ludwig Nohl.\par 
\vspace{12pt}
Also his letters to the Archduke Rudolph, Cardinal-Archbishop of Olm\"utz, K.W., from the collection of Dr.~Ludwig Ritter von K\"ochel.

\pagebreak

\mainmatter

\setcounter{chapter}{1}

\chapter{Life's Mission: 1815-1822}

\setcounter{section}{215}

\section{To Steiner and Co.}
The Adjutant's innocence is admitted, and there is an end of it!\par 
\vspace{12pt}
We beg you to be so good as to send uys two copies in score of the Symphony in A. We likewise wish to know when we may expect a copy of the Sonata
for Baroness von Ertmann, as she leaves this, most probably, the day after to-morrow. \par 
\vspace{12pt}
No. 3 -- I mean the enclosed note -- is from a musical friend in Silesia, not a rich man, for whom I have frequently had my scores written out. 
He wishes to have these works of Mozart in his library; as my servant, however, has the good fortune, by the grace of God, to be one of the 
greatest blockheads in the world (which is saying a good deal), I cannot make use of him for this purpose. Be so kind therefore as to send to Herr
-- (for the \textit{Generalissimus} can have no dealings with a petty tradesman), and desire him to \textit{write down the price of each work} and
send it to me with my two scores in A, and also and answer to my injunction about Ertmann, as early to-day as you can (\textit{presto, prestissimo}!) --
\textit{nota bene}, the \textit{finale} to be \textit{a march in double-quick time}. I recommend the best execution of these orders, so that no further
obstacle may intervene to my recovery. \par 
\vspace{12pt}
\begin{flushright}
    L. VAN BEETHOVEN, \\
    The best \textit{generalissimus} for the good, \\
    But the devil himself for the bad!
\end{flushright}

\section{To Steiner}
The Lieutenant-General is requested to send his \textit{Diabolum}, that I may tell him myself my opinion of the ``Battle,'' which is 
\textit{printed in the vilest manner}. There is much to be altered. \par 
\vspace{12pt}
\begin{flushright}
    THE G---S.
\end{flushright}

\section{To Tobias Haslinger}
MY GOOD ADJUTANT,-- \par 
\vspace{12pt}
Best of all little fellows! Do see again about that house, and get it for me. I am very anxious also to procure \textit{the treatise on education}.
It is of some importance to me to be able to compare my own opinions on this subject with those of others, and thus further improve them. As for our
juvenile Adjutant, I think I shall soon have hit on the right system for his education. Your \par 
\vspace{12pt}
\begin{flushright}
    CONTRA FA, \\
    \textit{Manu propria}.
\end{flushright}

\end{document}
